Investigating social change during cultural contact period using geometric morphometry of pottery shapes from Iron Age northeast Taiwan

### 2019-11-09 analysis/paper/paper.Rmd
update for the background section

Geometric morphometry methods (GMM) has been widely applied recently in archaeology for shape analysis, which explores morphological variability and similarity of archaeological materials to address the questions of anthropological interests. Different from linear measurement approach that captures shapes by measuring distances, angles or ratios of objects, geometric morphometrics methods use Cartesian coordinates of morphological structures to define the shapes. Landmarks, curves or outlines of objects can be transformed into coordinates by their unique location and orientation with respect to numerical values of coordinate axes [@Adams2004; @Lawing2010; @Slice2007]. According to different focuses and forms of original shape data, there are two common morphometric methods used to capture and analyze the shapes, landmark analysis and outline analysis [@Adams2004]. Landmark approaches assign a set of landmark points to specimen that can be specified on a coordinate system. The landmark coordinates are represented by numerical values with respect to defined coordinate axes. Based on the specific position of each point, objects are translated, rotated, and scaled into the same orientation and size using Procrustes superimposition technique to find the best fit between configuration shapes [@Lawing2010]. Landmarks

>>>>

Geometric morphometry methods (GMM) has been widely applied recently in archaeology for shape analysis, which explores morphological variability and similarity of archaeological materials to address the questions of anthropological interests. Different from linear measurement approach that captures shapes by measuring distances, angles or ratios of objects, geometric morphometrics methods use Cartesian coordinates of morphological structures to define the shapes [@Adams2004; @Lawing2010; @Slice2007]. Landmarks, curves or outlines of objects can be transformed into coordinates by their unique location and orientation with respect to numerical values of coordinate axes . According to different focuses and forms of original shape data, there are two common morphometric methods used to capture and analyze the shapes, landmark analysis and outline analysis [@Adams2004]. Landmark approaches assign a set of landmark points onto specimens that can be specified on a coordinate system. Each point has a coordinate represented by numerical values with respect to defined coordinate axes. Based on the distances and angles of interlandmarks, the objects are translated, rotated, and scaled into the same orientation and size using generalized Procrustes analysis (GPA). GPA method enables the comparison between shapes because it brings the landmark configurations of all objects into a common coordinate system, where differences are reflected based on landmark coordinate values [@Slice2007].
